---
title: "lin07Bmeshes"
author: "Erika Dona-JHS adaptation"
date: "2025-11-11"
output: html_document
---

To check if kn-unc4 intersection labels 07B neurons only in the segments that we predict (T1 and T2) we will save meshes of 07B neurons from manc.

#Set the environment
```{r setEnvironment, purl=TRUE}



#library(fancr)
##install.packages("remotes") # only one time 
#remotes::install_github("natverse/nat")# only one time 
#remotes::install_github("natverse/nat.templatebrains")# only one time 
library(nat)
library(nat.templatebrains)
library(tidyverse)
#1remotes::install_github("natverse/malevnc") #one time only
library(malevnc) 

#options(nat.cmtk.bindir = "C:/cygwin64/usr/local/bin")#point windows to CMTK location. not required for mac

options(nat.cmtk.bindir = "C:/cygwin64/usr/local/lib/cmtk/bin")
Sys.setenv(PATH = paste(
  "C:/cygwin64/bin",
  Sys.getenv("PATH"),
  sep = ";"
))

#test:
nat::cmtk.bindir()
cmtk.dof2mat(version = TRUE) #should say: "3.3.1"


#test
##list.files("C:/cygwin64/usr/local/lib/cmtk/bin")
#list.files("C:/cygwin64/bin", pattern = "cyg")
#Ensure cygwin isntalled these:
#file.exists("C:/cygwin64/bin/cygwin1.dll")
#file.exists("C:/cygwin64/bin/cyggomp-1.dll")
#cygwin installer: expand menue (not pendign but full; Install fftw3 3.3.10-1 Install libfftw3_3 and all related packegers to fftw (fouier transforms)

#library(wesanderson)
#library(ggpattern)
#library(ggbeeswarm)
#library(cowplot)
#library(dendextend)
#library(magick)


options(scipen = 999) # turns off scientific notation

outdir <- here::here("output", "test")
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
```

```{r}
VNCIS1=templatebrain("VNCIS1", dims = c(1024,768,163), voxdims = c(0.43,0.43,1), units = "microns")
MANC_MANCsym <- here::here("v1/Common/registrations/MANC_MANCsym.list")
MANCsym_VNCIS1 <- here::here("v1/Common/registrations/MANCsym_VNCIS1.list")

safe_read_mesh <- function(x) {
  tryCatch({
    malevnc::read_manc_meshes(x)
  }, error = function(e) {
    message("Skipping missing mesh: ", x)
    return(NULL)
  })
}

safe_read_neuron <- function(x) {
  tryCatch({
    malevnc::manc_read_neurons(x)
  }, error = function(e) {
    message("Skipping missing neuron: ", x)
    return(NULL)
  })
}

neurons_to_nrrds <- function(bodyids = c(16856),
                             refbrain = VNCIS1,
                             via = MANC_MANCsym,
                             reg = MANCsym_VNCIS1,
                             direction = "forward",
                             mesh = TRUE,
                             filename = "neurons_to_nrrds.nrrd") {
  
  if (mesh) {
    message("Reading meshes safely...")
    ns <- lapply(bodyids, safe_read_mesh)
    ns <- Filter(Negate(is.null), ns)
    ns <- as.neuronlist(unlist(ns, recursive = FALSE))
    ns <- ns / 1000
  } else {
    message("Reading neurons safely...")
    ns <- lapply(bodyids, safe_read_neuron)
    ns <- Filter(Negate(is.null), ns)
    ns <- as.neuronlist(unlist(ns, recursive = FALSE))
    ns <- ns / 125
  }
  
  if (length(ns) == 0) {
    stop("No valid neurons or meshes could be read. Aborting.")
  }
  
  if (!is.null(via)) {
    message("Applying via and final transformations...")
    ns.symm <- xform(x = ns, reg = via, direction = direction)
    ns.trans <- xform(x = ns.symm, reg = reg, direction = direction)
  } else {
    message("Applying single transformation...")
    ns.trans <- xform(x = ns, reg = reg, direction = direction)
  }
  
  im3 <- as.im3d(xyzmatrix(ns.trans), voxdims = refbrain)
  write.im3d(im3, file = filename)
}
```

```{r}
#get token 
Sys.setenv(NEUPRINT_TOKEN = "eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJlbWFpbCI6Imouc29mZmVyc0BnbWFpbC5jb20iLCJsZXZlbCI6Im5vYXV0aCIsImltYWdlLXVybCI6Imh0dHBzOi8vbGgzLmdvb2dsZXVzZXJjb250ZW50LmNvbS9hL0FDZzhvY0tRb1JBeUdrdjJmNF9UUU1LOTBIMEk4Nm5xLXN4eE04ZXhCMzEwYlFPaEdSdGNiQT1zOTYtYz9zej01MD9zej01MCIsImV4cCI6MTk1MzkzMDg4MH0.CV5AhDlg_RhVAwmBEdabcDpgXaY9kJ1qxhAj8PVOe3Q")

```


```{r load manc metadata}
# Load manc metadata
Sys.getenv("NEUPRINT_TOKEN")
malevnc::choose_malevnc_dataset("MANC")
options(malevnc.neuprint_dataset="manc:v1.2.3")
mancAnnotations <- malevnc::manc_neuprint_meta()
```

```{get4BcellID}

T2_04B <- mancAnnotations %>%
  filter(
    hemilineage == "04B",
    somaNeuromere == "T2",
    grepl("secondary", birthtime)
  )

cat("Secondary T2 04B neurons:", nrow(T2_04B), "\n")

T2_04B %>%
  select(bodyid, type, instance, hemilineage, somaNeuromere, birthtime) %>%
  print(n = Inf)




subset on 07B neurons
```{r}
outdir <- here::here("output", "test")
T107B <- mancAnnotations %>%
dplyr::filter(hemilineage == "07B", somaNeuromere == "T1")
neurons_to_nrrds(bodyids = T107B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc07BT1_meshes.nrrd"))
```
T2
```{r}
T207B <- mancAnnotations %>%
  filter(hemilineage == "07B", somaNeuromere == "T2")
neurons_to_nrrds(bodyids = T207B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc07BT2_meshes.nrrd"))
```
T3
```{r}
T307B <- mancAnnotations %>%
  filter(hemilineage == "07B", somaNeuromere == "T3")
neurons_to_nrrds(bodyids = T307B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc07BT3_meshes.nrrd"))
```

```{r}
A07B <- mancAnnotations %>%
  filter(hemilineage == "07B", grepl("A", somaNeuromere))
neurons_to_nrrds(bodyids = A07B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc07BA_meshes.nrrd"))
```

#12A
Also 12A is labelled by this intersection
```{r}
T112A <- mancAnnotations %>%
  filter(hemilineage == "12A", somaNeuromere == "T1")
neurons_to_nrrds(bodyids = T112A$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc12AT1_meshes.nrrd"))

T212A <- mancAnnotations %>%
  filter(hemilineage == "12A", somaNeuromere == "T2")
neurons_to_nrrds(bodyids = T212A$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc12AT2_meshes.nrrd"))
```
#18B
```{r}
T118B <- mancAnnotations %>%
  filter(hemilineage == "18B", somaNeuromere == "T1")
neurons_to_nrrds(bodyids = T118B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT118B_meshes.nrrd"))
                 
T218B <- mancAnnotations %>%
  filter(hemilineage == "18B", somaNeuromere == "T2")
neurons_to_nrrds(bodyids = T218B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT218B_meshes.nrrd"))

T318B <- mancAnnotations %>%
  filter(hemilineage == "18B", somaNeuromere == "T3")
neurons_to_nrrds(bodyids = T318B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT318B_meshes.nrrd"))                 
```

#23B
```{r}
T123B <- mancAnnotations %>%
  filter(hemilineage == "23B", somaNeuromere == "T1")
neurons_to_nrrds(bodyids = T123B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT123B_meshes.nrrd"))
T223B <- mancAnnotations %>%
  filter(hemilineage == "23B", somaNeuromere == "T2")
neurons_to_nrrds(bodyids = T223B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT223B_meshes.nrrd"))                
                 
T323B <- mancAnnotations %>%
  filter(hemilineage == "23B", somaNeuromere == "T3")
neurons_to_nrrds(bodyids = T323B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT323B_meshes.nrrd"))
```

#03B
Tk expression in secondary neurons should be specific for Abdominal segment on 03B neurons

```{r}
T103B <- mancAnnotations %>%
  filter(hemilineage == "03B", somaNeuromere == "T1", grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T103B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT103B_meshes.nrrd"))
T203B <- mancAnnotations %>%
  filter(hemilineage == "03B", somaNeuromere == "T2",grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T203B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT203B_meshes.nrrd"))                
                 
T303B <- mancAnnotations %>%
  filter(hemilineage == "03B", somaNeuromere == "T3",grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T303B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT303B_meshes.nrrd"))
A03B <- mancAnnotations %>%
  filter(hemilineage == "03B", grepl("A", somaNeuromere),grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = A03B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_manc03BA_meshes.nrrd"))                
```

#12B
```{r}
outdir <- "/Users/erikadona/Documents/Projects/RNAseq/NM_confocal_Jelly/meshes"
T1_12B <- mancAnnotations %>%
  filter(hemilineage == "12B", somaNeuromere == "T1", grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T1_12B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT1_12B_meshes.nrrd"))

T2_12B <- mancAnnotations %>%
  filter(hemilineage == "12B", somaNeuromere == "T2", grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T2_12B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT2_12B_meshes.nrrd"))

T3_12B <- mancAnnotations %>%
  filter(hemilineage == "12B", somaNeuromere == "T3", grepl("secondary", birthtime))
neurons_to_nrrds(bodyids = T3_12B$bodyid,
                 refbrain = VNCIS1,
                 via = MANC_MANCsym,
                 reg = MANCsym_VNCIS1,
                 direction = "forward",
                 mesh = TRUE,
                 filename = file.path(outdir, "VNCIS1_mancT3_12B_meshes.nrrd"))
```

